---
title: "UFOs in the UK"
output: 
  flexdashboard::flex_dashboard:
    theme:
      version: 4
      bg: "white"
      fg: "#4F556B" 
      primary: "#4C4C67"
      navbar-bg: "#4C4C67"
      base_font: 
        google: Prompt
      heading_font:
        google: Sen
    orientation: columns
    vertical_layout: fill
---

```{r setup, include=FALSE}
library(shiny)
library(flexdashboard)
library(scales)
library(tidyverse)
library(leaflet)
library(plotly)
library(gt)
library(gtExtras)
library(gtable)
library(tidytext)
library(textdata)
library(wordcloud)
library(patchwork)
library(ggtext)
library(fontawesome)
#load and preprocess data
data <- read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-06-25/ufo_sightings.csv")

ufo_palette <- c("#9FEE98", "#92D69D", "#85BDA2", "#7AA99D", "#6E9598", "#628387","#557076", "#515E6F", "#4F556B", "#4C4C67")

ufo_short_palette <- c("#9FEE98", "#7AA99D", "#4C4C67")

ufo_uk <- data %>% filter(country == "gb") %>%
  mutate(id = 1:nrow(.)) %>%
  mutate(ufo_shape = factor(ufo_shape))

```

Map of UFOs in the UK
=======================================================================


Column {data-width=650}
-----------------------------------------------------------------------

### Map of UFO encounters

```{r echo=F, message=F,warning=F}
ufo_shapes <- ufo_uk %>%
  mutate(ufo_shape = case_when(
    ufo_shape == "light" ~ "light",
    ufo_shape == "circle" ~ "circle",
    TRUE ~ "other"
  ))

factpal <- colorFactor(ufo_short_palette, ufo_shapes$ufo_shape)

ufo_tomap <- ufo_shapes %>%
  mutate(city_area = str_remove(city_area, "\\(.+\\)")) %>%
  group_by(city_area) %>%
  summarise(mean_enlen = mean(encounter_length, na.rm = T),
            mean_lat = mean(latitude),
            mean_lon = mean(longitude)) %>%
  inner_join(ufo_shapes %>%
               mutate(city_area = str_remove(city_area, "\\(.+\\)")) %>%
               group_by(city_area) %>%
               count(city_area))

ufo_icon <- makeIcon(
  iconUrl = "ufo_2.png",
  iconWidth = 20, iconHeight = 20,
  iconAnchorX = 1, iconAnchorY = 1
)

leaflet(ufo_tomap) %>% addTiles() %>%
  addMarkers(~mean_lon, ~mean_lat, icon = ufo_icon,
             label=~paste0("number of encounters: ",n, '\n; mean length: ', round(mean_enlen/60, 2), ' minutes'))



```



Column {data-width=350}
-----------------------------------------------------------------------

### Explanations

This is a flex dashboard analysing the ufo sightings in the UK in the XX Century.
The dashboard is divided into 3 parts:

- Map of ufo sightings along with number of encounters and their lengths
- Analysis of most common places and shapes of ufo sightings
- Analysis of descriptions of the ufo sightings

The dataset is taken from https://github.com/rfordatascience/tidytuesday/tree/master/data/2019/2019-06-25

### Map interpretation

The map on the left shows the places, 3 most common shapes shapes and lengths
of encounters in the UK.

- The most common place of UFO encounters is London with over 200 encounters
- There were far fewer encounters in the north of the UK than south

Places, shapes and time
=======================================================================

Column {data-width=500}
-----------------------------------------------------------------------

### Places


```{r echo=F, message = F, warning=F}
# this should have:
# 10 most common places
# 

ufo_10_places <- ufo_shapes %>%
  count(city_area) %>%
  arrange(desc(n)) %>%
  top_n(10) %>%
  mutate(city_area = str_remove(city_area, "\\(.+\\)"))%>%
  pull(city_area)

ufo_shapes %>%
  mutate(city_area = str_remove(city_area, "\\(.+\\)"))%>%
  filter(city_area %in% ufo_10_places) %>%
  group_by(city_area) %>%
  count(ufo_shape) %>%
  arrange(desc(n)) %>%
  summarise(list_data = list(n)) %>%
  inner_join(
    ufo_shapes %>%
      mutate(city_area = str_remove(city_area, "\\(.+\\)"))%>%
      filter(city_area %in% ufo_10_places) %>%
      group_by(city_area) %>%
      summarise(mean_length = round(mean(encounter_length/60),2))
  ) %>%
  gt() %>%
  gt_plt_bar_stack(column = list_data, palette = ufo_short_palette,labels = c("circle", "light", "other"), position = "stack")

```


### Box

```{r echo = F, message = F, warning = F}

year_max <- ufo_shapes %>%
  mutate(year = as.numeric(str_extract(date_time, '[0-9]{4}'))) %>%
  count(year) %>%
  arrange(desc(n)) %>%
  top_n(1)


valueBox(value = year_max$year,
         caption = paste0("Year with the most UFO encounters. ", year_max$n, " encounters were recorded that year."),
         icon = "fas fa-user-astronaut",
         color = "#7AA99D"
          )

```


Column {.tabset}
-----------------------------------------------------------------------

### Encounters across the years

```{r echo = F, message = F, warning = F}

timeline_plot <- ufo_shapes %>%
  mutate(year = as.numeric(str_extract(date_time, '[0-9]{4}'))) %>%
  count(year) %>%
  ggplot(aes(x = year, y = n, group = 1)) +
  geom_line() +
  scale_x_continuous(breaks = seq(1940, 2020, 10)) +
  theme_classic() +
  labs(title = "encounters surged in 2000s", x = "year", y = "number of encounters")

ggplotly(timeline_plot, tooltip = c("x", "y"))
```

### Places in time

```{r echo = F, message = F, warning = F}

places_time_plot <- ufo_shapes %>%
  mutate(city_area = str_remove(city_area, "\\(.+\\)"))%>%
  filter(city_area %in% ufo_10_places) %>%
  mutate(year = as.numeric(str_extract(date_time, '[0-9]{4}'))) %>%
  group_by(city_area) %>%
  count(year) %>%
  mutate(city_area = factor(city_area, ordered = TRUE,
                                  levels = c("coventry ", "edinburgh ", "leeds ",
                                             "liverpool ", "birmingham ", "bristol ",
                                             "manchester ", "nottingham ", "glasgow ",
                                             "london "))) %>%
  ggplot(aes(x = year, y = n, fill = city_area)) +
  geom_col(position = "stack") +
  scale_x_continuous(breaks = seq(1940, 2020, 10)) +
  scale_fill_manual(values = ufo_palette) +
  theme_classic() +
  theme(legend.position = "none") +
  labs(x = "year", y = "number of encounters", title = "Encounters in places across XX Century\nHover above the bar to get more information")
  
ggplotly(places_time_plot)

```


Descriptions
=======================================================================

Column {data-width=500}
-----------------------------------------------------------------------

### Wordcloud

100 most common words in the descriptions of the encounters:
```{r echo =F, warning=F, message=F}
data_desc <- ufo_uk %>%
  select(description, id)


data_token <- data_desc %>%
  unnest_tokens(output = "word", input = "description") %>%
  anti_join(stop_words)



desc_under_10_words <- data_token %>%
  count(id) %>%
  filter(n < 10) %>%
  pull(id)
  
data_token_filtered <- data_token %>%
  filter(!(id %in% desc_under_10_words))

data_token_countes <- data_token_filtered %>%
  count(word)


wordcloud(data_token_countes$word, data_token_countes$n, color = '#4C4C67',
          max.words = 100)

```


Most common word appearing in the descriptions is 44. Other common words are
related to collors, lights and shapes of the UFOs.
The emotions expressed in the descriptions were overally positive
with high levels of trust and anticipation.


Column {data-width=500}
-----------------------------------------------------------------------

### Emotions

```{r echo=F, message=F, warning=F}



nrc <- textdata::lexicon_nrc()

data_sent <- data_token_filtered %>%
  left_join(nrc, by = "word") %>%
  group_by(id) %>%
  count(sentiment) %>%
  mutate(percent = n/sum(n)) %>%
  drop_na()
  
data_sent2 <- data_sent %>%
  mutate(sentiment = factor(sentiment, ordered = T, levels = c("disgust", "anger", "surprise",
                                                   "sadness", "joy", "fear",
                                                   "negative", "trust", "anticipation",
                                                   "positive")),
         percent = round(percent,2))

plot_sent <- data_sent2 %>%
  ggplot(aes(x = sentiment, y = percent, color = sentiment)) +
  geom_point(position = position_jitter(), aes(text = paste0(
    "emotion: ", sentiment,
    "<br>",
    "percent:", scales::percent(percent)
  ))) +
  scale_color_manual(values = ufo_palette) +
  scale_y_continuous(labels = scales::percent) +
  coord_flip() +
  theme_classic() +
  theme(legend.position = "none") +
  labs(x = "", y = "percentage of text", title = "Most common emotions in descriptions were positive ones")

ggplotly(plot_sent, tooltip = c("text"))

```

